Cover letter for the original IJRR submission

### 2017-12-13 review_tro_2/reviewer_response.tex

\qt{The resulting polytope can contain a very large number of faces.  A last step is thus to simplify it using an incremental decimation method[38].
Variations of this method are commonly implemented in most authoring tools. For our experiments we used the Blender decimate tool. Details of its use can be found in the static webpage associated with this paper \url{http://stevetonneau.fr/files/publications/isrr15/tro_install.html\#decimate}.\\}

\section{Answers to Reviewer 1}
%~ \ndone reecriture p3 + enlever main video

\quot{First of all, I would like to thank the authors for the significant
improvements to the paper. The contributions are much clearer, the
algorithms are better described, and I believe that my initial concerns
about the paper have been addressed. In particular, I would like to
thank the authors for the detailed inline responses.}

\as{We in turn thank the reviewer, who significantly helped us improving the paper with her / his comments.}

@@

\as{Indeed, the sentence was somehow not removed from our latest version. We apologize for that, and guarantee that this time it was indeed removed!}

\leavevmode \\
\quot {
The dual reformulation of the LP in (11) in Sect. VI is fine, but is
a perfectly standard procedure (that, in fact, many LP solvers will
perform internally automatically). It would be sufficient to mention
that the dual problem is solved without giving its full formulation.
Whether the authors include the formulation is up to them.}

\as{Thanks for the comment. Indeed, using the dual formulation is quite common for LP solvers. However, after computing the dual in our case, it turns out that some of the constraints can be solved trivially, which allows to reduce the dimensionality of the problem, resulting in (11). We experimentally verified that this form is faster to solve with respect to the one from equation (10). For this reason, we believe that equation (11) is worth mentioning, and prefer to leave it, since it was given as an option by the reviewer. Finally, we removed the reference to the Slater's conditions since we agree with the reviewer that the procedure is standard.}

\leavevmode \\
\section{Answers to Reviewer 2}
\quot{All the issues I raised in the first review have been addressed in
the revised version. I'm satisfied with most of the answers and the
changes made to the paper. In particular the novelty of the paper with
respect to the conference works is clarified. I appreciate also the new
structure of the manuscript that integrates material previously
relegated to the Appendices into the main body. Now, this work seems to
be self contained enough for a journal. }
\as{Thanks for the comment. We agree with the reviewer that the paper is clearer this way, thanks to her / his comments.}

### 2017-12-13 body/appendix_algo.tex
ongoing reformating algo

\begin{algorithm}[!tbp]
\caption{Discretization of a path} \label{interpolate}
	\begin{algorithmic}[1]
	%~ \Function{GenerateConfiguration}{}
	\Function{Interpolate}{$s0$,$\mathbf{Q}^0$, $MAX\_TRIES$}
		\State $states \gets (s0)$ \Comment{List of states initialized with $s0$}
		\State $nb\_fail \gets 0$ 
		\State $i \gets 1;$ \Comment{Current index in the list}
		\While {$i < length(\mathbf{Q}^0)$}
			\State $pState \gets last\_element(states)$
			\State $s \gets$ \textsc{GenFullBody}$(pState, element(\mathbf{Q}^0,i))$
			\If {$s \not= 0 $}
				\State $nb\_fail \gets 0$
				\State $i \gets i+1$
				%~ \State \textbf{return} $\mathbf{q}^{0}$
			\Else
				\State $nb\_fail \gets nb\_fail + 1$
				%~ \If {$nb\_fail == MAX\_TRIES$}
					%~ \State \textbf{return} $FAILURE$
				%~ \EndIf				
				\State $s \gets $\textsc{IntermediateContactState}$(pState)$
				\If {$s == 0 \lor nb\_fail == MAX\_TRIES$}
					\State \textbf{return} $FAILURE$
				\EndIf		
			\EndIf
			\State $push\_back(states, s)$
		\EndWhile
		\State \textbf{return} $states$
	\EndFunction
\end{algorithmic}
\label{alg:interpolate}
\end{algorithm}

At each step, \textsc{GenFullBody} is called with the previous state as a parameter, as well
as a new root configuration. \textsc{GenFullBody} returns a new contact configuration, if it succeeded
in computing a configuration with only one contact switch occurring.
Otherwise, the method \textsc{IntermediateContactState} is called.
It repositions one end effector (either a free limb, or the oldest active contact) towards a new contact position if possible.
This repositioning allows to increase the odds that the contact can be maintained at the next step.
The method \textit{length} returns the length of a list, \textit{last\_element} returns the last element
of a list, \textit{element} returns the element contained by a list at a given index, and \textit{push\_back} inserts an element at the end of a list.
Algorithm~\ref{alg:pi} gives the pseudo code for \textsc{GenFullBody}.

\begin{algorithm}[!tbp]
\caption{Full body contact generation method} \label{interpolate}
	\begin{algorithmic}[1]
	%~ \Function{GenerateConfiguration}{}
	\Function{\textsc{GenFullBody}}{$pState$,$\mathbf{q}^0$}
		\State $newState \gets CreateState(\mathbf{q}^0, ContactLimbs(pState))$
		%~ \State /*First try to maintain previous contacts*/
		\State $nConBroken \gets 0$
		\For {\textbf{each} $k$ in $ContactLimbs(pState)$}
			\If {$!$\textsc{MaintainContact}$(pState,\mathbf{q}^0,k)$}
				\State $nConBroken \gets nConBroken +1$
				\If {$nConBroken > 1$}				
					\State \textbf{return} $0$
				\EndIf				
				\State $MarkFree(newState,k)$
			\Else 					
				\State $MarkContact(newState,k)$
			\EndIf
		\EndFor
		\For {\textbf{each} $k$ in $FreeLimbs(newState)$}
			\If {\textsc{GenerateContact}$(newState,k)$}	
				\State $MarkContact(newState,k)$
				\State \textbf{return} $newState$
			\EndIf
		\EndFor
		\If {\textsc{IsInStaticEquilibrium}$(newState)$}
			\State \textbf{return} $newState$
		\Else
			\State \textbf{return} $0$
		\EndIf
	\EndFunction
\end{algorithmic}
\label{alg:pi}
\end{algorithm}

The method \textsc{MaintainContact}$(pState,\mathbf{q}^0,k)$ performs inverse kinematics to reach the previous contact position for the Limb.
If it succeeds, the new limb configuration is assigned to $k$. If it fails, a random collision free configuration is assigned to $k$.

The method \textsc{IsInStaticEquilibrium} returns whether a given state is in static equilibrium. $CreateState$ creates a new state, given
a root configuration and the list of active contacts.
$ContactLimbs$ (respectively $FreeLimbs$) returns the list of limbs that are in contact (respectively \textit{not} in contact).  $MarkContact$ (respectively $MarkFree$)  marks a limb as in contact (respectively \textit{not} in contact). These methods follow a FIFO approach: the first Limb chronologically marked as in contact (respectively not in contact) is returned first by $ContactLimbs$ (respectively $FreeLimbs$). This allows the algorithm to be deterministic even though it can handle acyclic motions.
%~ $add\_contact\_limb(newState,k)$ marks a limb as in contact at a given state.

The pseudo code for the method \textsc{IntermediateContactState} is given by Algorithm~\ref{alg:repo}.

\textsc{GenerateContact}$(state,k)$ is a call to the contact generator presented in the following Section~\ref{sec:single_contact}.
It generates a contact configuration in static equilibrium, and assigns the corresponding configuration of the Limb $k$ to the state configuration.
If it fails, $k$ remains unchanged if it is collision free, otherwise it is assigned a random collision free configuration.

>>>>

\begin{algorithm}[!tbp]
\caption{Discretization of a path} \label{interpolate}
	\begin{algorithmic}[1]
	%~ \Function{GenerateConfiguration}{}
	\Function{Interpolate}{$s0$,$\mathbf{Q}^0$, $MAX\_TRIES$}
		\State $states \gets (s0)$ \Comment{List of states initialized with $s0$}
		\State $nb\_fail \gets 0$ 
		\State $i \gets 1;$ \Comment{Current index in the list}
		\While {$i < length(\mathbf{Q}^0)$}
			\State $pState \gets lastElement(states)$
			\State $s \gets$ \textsc{GenFullBody}$(pState, element(\mathbf{Q}^0,i))$
			\If {$s \not= 0 $}
				\State $nb\_fail \gets 0$
				\State $i \gets i+1$
				%~ \State \textbf{return} $\mathbf{q}^{0}$
			\Else
				\State $nb\_fail \gets nb\_fail + 1$
				%~ \If {$nb\_fail == MAX\_TRIES$}
					%~ \State \textbf{return} $FAILURE$
				%~ \EndIf				
				\State $s \gets $\textsc{IntermediateContactState}$(pState)$
				\If {$s == 0 \lor nb\_fail == MAX\_TRIES$}
					\State \textbf{return} $FAILURE$
				\EndIf		
			\EndIf
			\State $pushBack(states, s)$
		\EndWhile
		\State \textbf{return} $states$
	\EndFunction
\end{algorithmic}
\label{alg:interpolate}
\end{algorithm}

At each step, \textsc{GenFullBody} (Algorithm~\ref{alg:pi}) is called with the previous state as a parameter, as well
as a new root configuration. \textsc{GenFullBody} returns a new contact configuration, if it succeeded
in computing a configuration with only one contact switch occurring.
Otherwise, the method \textsc{IntermediateContactState} (Algorithm~\ref{alg:repo}) is called.
It repositions one end effector (either a free limb, or the oldest active contact) towards a new contact position if possible.
This repositioning allows to increase the odds that the contact can be maintained at the next step.
The method \textit{length} returns the length of a list, \textit{lastElement} returns the last element
of a list, \textit{element} returns the element contained by a list at a given index, and \textit{pushBack} inserts an element at the end of a list.
%~ Algorithm~\ref{alg:pi} gives the pseudo code for \textsc{GenFullBody}.

\begin{algorithm}[!tbp]
\caption{Full body contact generation method} \label{interpolate}
	\begin{algorithmic}[1]
	%~ \Function{GenerateConfiguration}{}
	\Function{\textsc{GenFullBody}}{$pState$,$\mathbf{q}^0$}
		\State $newState \gets CreateState(\mathbf{q}^0,$ 
        \State $\quad \quad \quad \quad  \quad  \quad  \quad  \quad  \quad \quad  \quad ContactLimbs(pState))$
		%~ \State /*First try to maintain previous contacts*/
		\State $nConBroken \gets 0$
		\For {\textbf{each} $k$ in $ContactLimbs(pState)$}
			\If {$\neg$\textsc{MaintainContact}$(pState,\mathbf{q}^0,k)$}
				\State $nConBroken \gets nConBroken +1$
				\If {$nConBroken > 1$}				
					\State \textbf{return} $0$
				\EndIf				
				\State $MarkFree(newState,k)$
			\Else 					
				\State $MarkContact(newState,k)$
			\EndIf
		\EndFor
		\For {\textbf{each} $k$ in $FreeLimbs(newState)$}
			\If {\textsc{GenerateContact}$(newState,k)$}	
				\State $MarkContact(newState,k)$
				\State \textbf{return} $newState$
			\EndIf
		\EndFor
		\If {$IsInStaticEquilibrium(newState)$}
			\State \textbf{return} $newState$
		\Else
			\State \textbf{return} $0$
		\EndIf
	\EndFunction
\end{algorithmic}
\label{alg:pi}
\end{algorithm}

The procedure \textsc{MaintainContact}$(pState,\mathbf{q}^0,k)$ performs inverse kinematics to reach the previous contact position for the limb.
If it succeeds, the new limb configuration is assigned to $k$. If it fails, a random collision free configuration is assigned to $k$.

The method $IsInStaticEquilibrium$ returns whether a given state is in static equilibrium. $CreateState$ creates a new state, given
a root configuration and the list of active contacts.
$ContactLimbs$ (respectively $FreeLimbs$) returns the list of limbs that are in contact (respectively \textit{not} in contact).  $MarkContact$ (respectively $MarkFree$)  marks a limb as in contact (respectively \textit{not} in contact). These methods follow a FIFO approach: the first limb chronologically marked as in contact (respectively not in contact) is returned first by $ContactLimbs$ (respectively $FreeLimbs$). This allows the algorithm to be deterministic even though it can handle acyclic motions.
%~ $add\_contact\_limb(newState,k)$ marks a limb as in contact at a given state.

%~ The pseudo code for the method \textsc{IntermediateContactState} is given by Algorithm~\ref{alg:repo}.

\textsc{GenerateContact}$(state,k)$ is a call to the contact generator presented in the following Section~\ref{sec:single_contact}.
It generates a contact configuration in static equilibrium, and assigns the corresponding configuration of the limb $k$ to the state configuration.
If it fails, $k$ remains unchanged if it is collision free, otherwise it is assigned a random collision free configuration.
